---
title: "Project Two"
output:
  pdf_document: default
  html_document: default
---

Due Oct. 21 at 11:59 PM. 

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r}
surveys_complete <- read_csv("Data Output/surveys_complete.csv")
```

```
Predictor is weight and response is hindfoot_length
```

2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
library(ggplot2)
ggplot(data = data, mapping = aes(x = weight, y = hindfoot_length)) + geom_point(size = 0.5) +
  geom_smooth(method = "lm", color = "navy", size = 1.0, fill = "deeppink") +
  labs(x = "weight", y = "hindfoot_length") +
  annotate("text", x = 20, y = 70, label = "R^2 == 0.4673", parse = T, size = 5) +
  theme(text = element_text(size = 20))
```

```
Yes, the data appears to be related linearly.
```


3) Fit the linear model. View the summary. (5 pts)


```{r}
linear_model <- lm(hindfoot_length ~ weight, data = data)
summary(linear_model)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
Yes, the summary makes sense when it is compared to the plot. Based on the data, under the coefficients I would say yes, our model has good predictive power.Yes, I believe the predictor predicts the response based on the residual standard error, intercept and R-squared as their data backs up the data in the graph.
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```
ggplot(data = data, mapping = aes(x = weight, y = hindfoot_length)) + geom_point(size = 0.5) +
 geom_smooth(method = "lm", color = "navy", size = 1.0, fill = "deeppink") +
 labs(x = "weight, y = "hindfoot_length") +
 annotate("text", x = 20, y = 70, label = "R^2 ==0.4673", parse = T, size = 5) + theme(text = element_text(size = 20))
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}

broom::augment(linear_model) -> augmented_fit
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col = "purple")
```

Why is normality of residuals important? 

```{r}
I believe the normality of residuals is important because it is a assumption of a linear model. Which means if the residuals are normal, the assumption is valid.
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
ggplot(data = data, mapping = aes(x = weight, y = hindfoot_length)) + geom_point(size = 0.5) + 
  facet_wrap(facets = vars(genus))

Yes, there is an interspecific variation.
```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}
library(ggplot2)
ggplot(data = data, aes(x = sex, y = weight, color = sex)) + geom_jitter(size = 0.5) +
  geom_smooth(method = "lm", color = "navy", size = 1.0, fill = "deeppink4") +
  labs(x = "sex", y = "weight") +
  annotate("text", x = 20, y = 70, label = "R^2 == 0.4673", parse = T, size = 5) +
  theme(text = element_text(size = 20))
```

2) Try an ANOVA of this data (5 pt)

```{r}
aov(linear_model) -> anova_model_fit
summary(anova_model_fit)
```

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)


```{r}
model_fit <- lm(hindfoot_length ~ weight, data = data)
summary(model_fit)
```

```
A linear model is used to predict a continuous outcome on the basis of one or more continuous predictor variables. Whereas, ANOVA is used to predict a continuous outcome on the basis of one or more categorical predictor variables.
```

3) Plot the lm with the data, with points colored by sex. (10 pts)


```{r}
library(ggplot2)
ggplot(data = data, aes(x = sex, y = weight, color = sex)) + geom_jitter(size = 0.5) + 
geom_smooth(method = "lm")
  
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
model_fit <- lm(weight ~ species_id, data = data)
summary(model_fit)
```

```{r}
library(ggplot2)
ggplot(data = data, aes(x = weight, y = species_id, color = sex)) +
  geom_jitter(size = 3.5) + geom_smooth(method = "lm", color = "navy", size = 1.0, fill = "deeppink4") +
  labs(x = "sex", y = "weight") +
  annotate("text", x = 100, y = 50, label = "R^2 == 7.351e-05", parse = T, size = 5) +
  theme(text = element_text(size = 20))
```

```
After plotting the data, my interpretation definitely has changed from my results the weight vs. species_id differs tremendously, but it also shows how each species weight is proportional to what type of species is being displayed.
```

## Part Three


1) Add and commit this document (5 pts)

```
install.packages("use.this")
library(usethis)
  git config --global user.name "John Williams"
  git config --global user.email "john.williams-8@selu.edu"
use_git()
Selection: 2
Go to git hub tab
Select project two from files
Check the box next to it
Hit commit
Type in a commit message, "Project Two upload for class/project"
Hit commit
>>> /usr/bin/git commit -F /tmp/RtmpPBBNVi/git-commit-message-e276d3e71f.txt
[Master 87d9358] Project Two Upload for project two/class
 2 files changed, 332 insertions(+)
 create mode 100644 Projects/Project_Two.Rmd
 create mode 100644 Projects/Project_two/Project_TwoWilliams.Rmd
```

2) Push your changes to github (10 pts)

```
install.packages("use.this.")
library(usethis)
  git config --global user.name "John Williams"
  git config --global user.email "john.williams-8@selu.edu"
use_git()
Selection: 2
create_github_token()
gitcreds::gitcreds_set()
Go to git hub tab
Select project two from files
Check the box next to it
Click Git push

"I am finished with my project".
```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
